---
title: "Distribution Gap Report — `r params$taxon`"
subtitle: "Candidate new Brazilian state records, from occurrence evidence absent from the Flora e Funga do Brasil distribution"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
author: "Generated by [floraR](https://dboslab.github.io/floraR-website/)"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: false
    df_print: paged
    css: !expr system.file("rmd/flora_report.css", package = "floraR")
params:
  data_path: NULL
  taxon: "TAXON"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  out.width = "100%"
)

library(DT)

# Load pre-computed analysis data
d <- readRDS(params$data_path)
taxon <- d$taxon
comma <- function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)
logo_path <- system.file("figures", "floraR_hex_sticker.png", package = "floraR")
```

<div style="float:right; margin-top:-80px; margin-left:15px;">
  <a href="https://dboslab.github.io/floraR-website/" target="_blank">
    <img src="`r logo_path`"
         style="width:120px;transition:opacity 0.2s ease;"
         onmouseover="this.style.opacity='0.8'"
         onmouseout="this.style.opacity='1'"
         alt="floraR package logo">
  </a>
</div>

---

## Overview

Brazilian states with occurrence evidence for **`r taxon`** from
`r paste(setdiff(d$source_summary$Source, "FFB (official)"), collapse = ", ")` were compared
against the states officially listed for this taxon in the
[Flora e Funga do Brasil (FFB)](https://floradobrasil.jbrj.gov.br/consulta/) distribution,
to flag **candidate new state records**.
`r if (!identical(d$ffb_name, d$taxon)) paste0("In FFB, ", d$taxon, " is a synonym of *", d$ffb_name, "*, whose distribution was used.") else ""`
`r if (!isTRUE(d$ffb_found)) paste0("**", d$taxon, " was not found in the current FFB checklist**, so every state with records is flagged.") else if (d$n_in_ffb == 0) paste0("**FFB lists no Brazilian state for *", d$ffb_name, "*** (as for some marine algae), so every state with records is flagged.") else ""`
`r if (!is.na(d$gbif_name)) paste0("GBIF records were matched to *", d$gbif_name, "*, including its synonyms and infraspecific taxa.") else ""`

Occurrence records are evidence to review, not proof: a record from a new state may be a
cultivated plant, a misidentification, or a georeferencing error. The individual records
behind each candidate state are listed at the end of this report.

```{r kpi-boxes}
htmltools::tags$div(
  class = "kpi-row",
  htmltools::tags$div(
    class = "kpi-box kpi-green",
    htmltools::tags$div(class = "kpi-value", comma(d$n_in_ffb)),
    htmltools::tags$div(class = "kpi-label", "States already in FFB distribution")
  ),
  htmltools::tags$div(
    class = "kpi-box kpi-gold",
    htmltools::tags$div(class = "kpi-value", comma(d$n_states_total)),
    htmltools::tags$div(class = "kpi-label", "States with any evidence")
  ),
  htmltools::tags$div(
    class = "kpi-box kpi-red",
    htmltools::tags$div(class = "kpi-value", comma(d$n_new_candidates)),
    htmltools::tags$div(class = "kpi-label", "New state record candidates")
  )
)
```

---

## Evidence by Source

```{r table-source}
src <- d$source_summary

DT::datatable(
  src,
  colnames = c("Source", "States with evidence", "Records"),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Table 1. Number of states and records with evidence, by source."
  ),
  options = list(pageLength = 5, dom = "t")
) %>%
  DT::formatStyle("States",
                  background = DT::styleColorBar(src$States, "#EEF6EC"),
                  backgroundSize = "98% 88%",
                  backgroundRepeat = "no-repeat",
                  backgroundPosition = "center")
```

---

## Complete State-by-State Table

The table below lists every Brazilian state with evidence for **`r taxon`**, flagging
`New_state_record_candidate = TRUE` for states with occurrence evidence that are not yet
in FFB's official distribution. Use the column filters to narrow the list, and the
buttons above the table to copy or download it as CSV/Excel.

```{r table-states}
tbl <- d$result

DT::datatable(
  tbl,
  rownames = FALSE,
  filter = list(position = "top", clear = TRUE, plain = FALSE),
  extensions = c("Buttons"),
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    paste0("Table 2. All ", comma(nrow(tbl)),
           " state(s) with evidence for ", taxon, ".")
  ),
  options = list(
    pageLength = 30,
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    scrollX = TRUE,
    search = list(regex = TRUE, caseInsensitive = TRUE)
  )
) %>%
  DT::formatStyle(
    "New_state_record_candidate",
    backgroundColor = DT::styleEqual(c(TRUE, FALSE), c("#fde8ea", "#EEF6EC"))
  )
```

---

## Individual Records for the New State Candidates

Unlike the state-level table above, this table lists the individual occurrence records
found in Brazilian states that are not yet in FFB's official distribution for
**`r taxon`** (up to 50 GBIF records per state), each linking to the record in its
source database.

```{r table-records}
recs <- d$records_detail

if (is.null(recs) || nrow(recs) == 0) {
  htmltools::p(
    "No individual records were retrieved for the new-state-record candidates ",
    "(there may be no candidates, or no record-level source among 'sources').",
    style = "color: #6c757d; font-style: italic;"
  )
} else {
  recs$RecordLink <- ifelse(
    !is.na(recs$RecordLink) & grepl("^https?://", recs$RecordLink),
    paste0('<a href="', recs$RecordLink, '" target="_blank">', recs$RecordLink, '</a>'),
    ifelse(is.na(recs$RecordLink), "", recs$RecordLink)
  )

  DT::datatable(
    recs,
    rownames = FALSE,
    escape = FALSE,
    filter = list(position = "top", clear = TRUE, plain = FALSE),
    extensions = c("Buttons"),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold;",
      paste0("Table 3. ", comma(nrow(recs)),
             " individual record(s) from the new-state-record candidates for ", taxon, ".")
    ),
    options = list(
      pageLength = 20,
      dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollX = TRUE,
      search = list(regex = TRUE, caseInsensitive = TRUE),
      lengthMenu = list(c(10, 20, 50, 100, -1),
                        c("10", "20", "50", "100", "All"))
    )
  )
}
```

---

<footer style="text-align:center; color:#adb5bd; font-size:0.85em; margin-top:2em; padding-top:1em; border-top:1px solid #dee2e6;">
Analysis produced by <strong>floraR</strong> &bull;
Flora e Funga do Brasil &bull;
`r paste(setdiff(d$source_summary$Source, "FFB (official)"), collapse = " &bull; ")` &bull;
`r format(Sys.Date(), "%Y")`
</footer>
