---
title: "Species Gap Report — `r params$taxon`"
subtitle: "Species recorded in Brazil that are missing from the Flora e Funga do Brasil checklist"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
author: "Generated by [floraR](https://dboslab.github.io/floraR-website/)"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: false
    df_print: paged
    css: !expr system.file("rmd/flora_report.css", package = "floraR")
params:
  data_path: NULL
  taxon: "TAXON"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  out.width = "100%"
)

library(DT)

# Load pre-computed analysis data
d <- readRDS(params$data_path)
taxon <- d$taxon
comma <- function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)
logo_path <- system.file("figures", "floraR_hex_sticker.png", package = "floraR")
```

<div style="float:right; margin-top:-80px; margin-left:15px;">
  <a href="https://dboslab.github.io/floraR-website/" target="_blank">
    <img src="`r logo_path`"
         style="width:120px;transition:opacity 0.2s ease;"
         onmouseover="this.style.opacity='0.8'"
         onmouseout="this.style.opacity='1'"
         alt="floraR package logo">
  </a>
</div>

---

## Overview

Species-level names in **`r taxon`** (genus) were compared against the current
[Flora e Funga do Brasil (FFB)](https://floradobrasil.jbrj.gov.br/consulta/) checklist,
using `r if (length(d$sources) == 2) "two independent kinds of evidence that a species occurs in Brazil" else "the following evidence that a species occurs in Brazil"`:

`r paste0("- ", c(if ("POWO" %in% d$sources) "**POWO distributions** - species that [Plants of the World Online](https://powo.science.kew.org) maps in Brazil (the TDWG areas Brazil North, Northeast, West-Central, Southeast, and South);", if ("GBIF" %in% d$sources) "**GBIF type specimens** - species whose type specimens were collected in Brazil, as published by herbaria through [GBIF](https://www.gbif.org)."), collapse = "\n")`

Names are compared by their *accepted* name as well as the name itself, since checklists
disagree on which genus a species belongs to, and orthographic variants (case, hyphens,
*-ii*/*-i* endings) are treated as the same name. Each name below is a candidate for
inclusion in the FFB monograph of its group; `r if (length(d$sources) == 2) "names reported by **both** sources are listed first, as the strongest candidates." else "the table lists the evidence behind each one."`
`r if ("POWO" %in% d$sources && !isTRUE(d$brazil_filtered)) "POWO names were not filtered by distribution: names POWO places only outside Brazil are included for manual review (see the Brazil_evidence column)." else ""`

**Data sources:** `r paste(d$source_lines, collapse = "; ")`.

```{r kpi-boxes}
box <- function(cls, value, label) {
  htmltools::tags$div(class = paste("kpi-box", cls),
                      htmltools::tags$div(class = "kpi-value", comma(value)),
                      htmltools::tags$div(class = "kpi-label", label))
}
kpi_boxes <- list(box("kpi-green", d$n_in_ffb, "Species of the genus registered in FFB"))
if (!is.na(d$n_powo_names)) {
  kpi_boxes <- c(kpi_boxes, list(box("kpi-gold", d$n_powo_names,
                                     paste0("Species-level names in POWO for ", taxon))))
}
if (!is.na(d$n_gbif_names)) {
  kpi_boxes <- c(kpi_boxes, list(box("kpi-gold", d$n_gbif_names,
                                     "Names with Brazilian type specimens in GBIF")))
}
kpi_boxes <- c(kpi_boxes, list(box("kpi-red", d$n_missing, "Missing from FFB")))
if (length(d$sources) == 2) {
  kpi_boxes <- c(kpi_boxes, list(box("kpi-amber", d$n_both, "Reported by both POWO and GBIF")))
}
do.call(htmltools::tags$div, c(class = "kpi-row", kpi_boxes))
```

---

## Species Missing from FFB

The table below lists all `r comma(d$n_missing)` species-level names in **`r taxon`**
that are not currently in FFB, with the evidence from each source and links to each
name's pages. Use the column filters to narrow the list - e.g. `Found_in` - and the
buttons above the table to copy or download it as CSV/Excel.

```{r table-missing}
tbl <- d$result

for (col in intersect(c("POWO_URL", "IPNI_URL", "GBIF_URL"), names(tbl))) {
  tbl[[col]] <- ifelse(
    !is.na(tbl[[col]]) & nzchar(tbl[[col]]),
    paste0('<a href="', tbl[[col]], '" target="_blank">', tbl[[col]], '</a>'),
    ""
  )
}

dt <- DT::datatable(
  tbl,
  rownames = FALSE,
  escape = FALSE,
  filter = list(position = "top", clear = TRUE, plain = FALSE),
  extensions = c("Buttons"),
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    paste0("Table 1. All ", comma(d$n_missing), " species of ", taxon,
           " missing from FFB.")
  ),
  options = list(
    pageLength = 20,
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    scrollX = TRUE,
    search = list(regex = TRUE, caseInsensitive = TRUE),
    lengthMenu = list(c(10, 20, 50, 100, -1),
                      c("10", "20", "50", "100", "All"))
  )
)
if (length(d$sources) == 2) {
  dt <- DT::formatStyle(dt, "Found_in",
                        backgroundColor = DT::styleEqual("POWO; GBIF", "#EEF6EC"),
                        fontWeight = DT::styleEqual("POWO; GBIF", "bold"))
}
dt
```

---

<footer style="text-align:center; color:#adb5bd; font-size:0.85em; margin-top:2em; padding-top:1em; border-top:1px solid #dee2e6;">
Analysis produced by <strong>floraR</strong> &bull;
Flora e Funga do Brasil &bull;
`r paste(d$sources, collapse = " &bull; ")` &bull;
`r format(Sys.Date(), "%Y")`
</footer>
